DeepSphere: a graph-based spherical CNN

### 2019-09-25 main.tex
Updates from Overleaf

\begin{abstract}
Selecting the right convolution for a spherical neural network is a delicate tradeoff between efficiency and  exactitude with respect to equivariance. DeepSphere, as a graph based method strikes as a controllable balance between these two desiderata. 
% Say something here about isotropy?
This contribution (i) studies empirically and theoretically how the construction of its underlying graph affects equivariance, and (ii) evaluates its performance on a large set of relevant problems.

Comparison with previous work on shape classification and climate event detection suggests that anisotropic filters are an unnecessary price to pay.
% memory, computations, implementation complexity
\end{abstract}

@@

\Tabref{shrec17_retrieval} shows the results obtained from the competition's official evaluation script.
\Tabref{modelnet40} shows results on ModelNet40.

@@

\begin{table}
    \centering
    \begin{tabular}{l cccc}
	\toprule
    & no/no & z/z & SO3/SO3 & z/SO3 \\
	\midrule
	\citet{cohen2018sphericalcnn} & 85.0 & - & - & - \\
	\citet{jiang2019sphericalcnn} & 90.5 & - & - & - \\
	\citet{esteves2018sphericalcnn} & - & 88.9 & 86.9 & 76.7 \\
	%Esteves \emph{MVCNN} & 94.69 & - & - & - \\
	DeepSphere & 87.8 & 86.8 & 86.7 & 76.9 \\
	\bottomrule
    \end{tabular}
    \caption{
		Accuracy results on ModelNet40.
		``no'' denotes no augmentation, ``z'' is augmented with rotations around the Z-axis, and ``SO3'' with ZYZ rotations.
		\todo{The first is training time augmentation, the second test time augmentation.}
	}
    \label{tab:modelnet40}
\end{table}

@@

\begin{table}
    \centering
	\begin{tabular}{l l l l l}
		\toprule
        & TC & AR & BG & mean \\
		\midrule
		\cite{mudigonda2017climateevents} & 74 & 65 & 97 & 78.67 \\
		\cite{jiang2019sphericalcnn} (paper) & 94 & 93 & 97 & 94.67 \\
		\cite{jiang2019sphericalcnn} (rerun) & 93.9 & 95.7 & 95.2 & 94.95 \\
        \cite{cohen2019gauge} (S2R) & 97.8 & 97.3 & 97.3 & 97.5 \\
        \cite{cohen2019gauge} (R2R) & 97.9 & 97.8 & 97.4 & 97.7 \\
		\midrule
		DS (Jiang architecture, weighted loss) & 97.1 & 97.6 & 96.5 & 97.1 \\
        DS (weighted loss) & $97.4\pm 1.1$ & $97.7\pm 0.7$ & $98.2\pm 0.5$ & $97.8\pm 0.3$ \\
		DS (wider architecture, weighted loss) & 91.5 & 93.4 & 99.0 & 94.6 \\
		\midrule
        DS (Jiang architecture, non-weighted loss) & 33.6 & 93.6 & 99.3 & 75.5 \\
        DS (non-weighted loss) & $69.2\pm 3.7$ & $94.5\pm 2.9$ & $99.7\pm 0.1$ & $87.8\pm 0.5$ \\
        DS (wider architecture, non-weighted loss) & 73.4 & 92.7 & 99.8 & 88.7 \\
% 		\midrule
% 		\todo{DS (Cohen architecture, weighted loss)} & 95.5 & 96.9 & 93.9 & 95.4 \\
% 		\todo{DS (Cohen architecture, non-weighted loss)} & 56.4 & 90.0 & 98.9 & 81.8 \\
% 		\todo{DS (full dataset, non-weighted loss)} & 31.3 & 75.2 & 99.9 & 68.80 \\
		\bottomrule
    \end{tabular}
    \caption{
		Results on climate event segmentation: accuracy.
		Tropical cyclones (TC) and atmospheric rivers (AR) are the two positive classes, against the background (BG).
		\todo{\citet{mudigonda2017climateevents} is not directly comparable as they don't use the same input feature maps.}
		Note that a non-weighted cross-entropy loss is not optimal for the accuracy metric.
		\label{tab:climate:accuracy}
	}
\end{table}

@@

\begin{table}
	\centering
	\begin{tabular}{l r r r}
		\toprule
		& \multicolumn{1}{c}{size} & \multicolumn{2}{c}{speed} \\
        \cmidrule(lr){2-2} \cmidrule(lr){3-4}
		& params & inference & training \\
		\midrule
        %Mudigonda et al. & 97 & 74 & 65 & 78.67 & - & - & - \\
		\cite{jiang2019sphericalcnn} & 330\,k & 10\,ms & 10\,h \\ % 328'339
		% \cite{cohen2019gauge} (S2R) & - & - & - \\
        % \cite{cohen2019gauge} (R2R) & - & - & - \\
		DeepSphere (Jiang architecture) & 590\,k & 5\,ms & 3\,h \\ % 590k
        DeepSphere & 13\,M & 33\,ms & 13\,h \\ % 12'926'432
		DeepSphere (wider architecture) & 52\,M & 50\,ms & 20\,h \\
% 		\todo{DS (Cohen architecture)} & 2\,M & 11\,ms & 4\,h \\
% 		\todo{DS (full dataset)} & - & 567\,ms & 48\,h \\
		\bottomrule
    \end{tabular}
    \caption{
		Results on climate event segmentation: size and speed.
	}
		\label{tab:climate:speed}
\end{table}

@@

concat is the operation that concatenate the results of the corresponding encoder layer.
% Not similar to Jiang and 4 times the feature maps.

@@

% DeepSphere with a wider architecture is exactly the same, but with 2 times the number of feature maps.
% \subsubsection*{DS (Jiang architecture)}

DeepSphere with Jiang architecture\\
encoder:\\
\begin{dmath}
    [GC_{8}\, +\, BN\, +\, ReLU]_{L5}\,+\,\textrm{Pool}\,+\, [GC_{16}\, +\, BN\, +\, ReLU]_{L4}\, +\, \textrm{Pool}\, +\, [GC_{32}\, +\, BN\, +\, ReLU]_{L3}\, +\, \textrm{Pool}\, +\, [GC_{64}\, +\, BN\, +\, ReLU]_{L2}\, +\,\textrm{Pool}\, +\, [GC_{128}\, +\, BN\, +\, ReLU]_{L1}\, +\, \textrm{Pool}\,  +\, [GC_{128}\, +\, BN\, +\, ReLU]_{L0}
\end{dmath}
decoder:\\
\begin{dmath}
    \textrm{Unpool}\, +\,[GC_{128}\, +\, BN\, +\, ReLU]_{L1}\, +\, \textrm{concat}\, +\, [GC_{128}\, +\, BN\, +\, ReLU]_{L1}\, +\, \textrm{Unpool}\, +\, [GC_{64}\, +\, BN\, +\, ReLU]_{L2}\, +\, \textrm{concat}\, +\, [GC_{64}\, +\, BN\, +\, ReLU]_{L2}\, +\, \textrm{Unpool}\, +\, [GC_{32}\, +\, BN\, +\, ReLU]_{L3}\, +\, \textrm{concat}\, +\, [GC_{32}\, +\, BN\, +\, ReLU]_{L3}\, +\,\textrm{Unpool}\, +\, [GC_{16}\, +\, BN\, +\, ReLU]_{L4}\,+\, \textrm{concat}\, +\, [GC_{16}\, +\, BN\, +\, ReLU]_{L4}\, +\,\textrm{Unpool}\,  +\, [GC_{8}\, +\, BN\, +\, ReLU]_{L5}\,+\,\textrm{concat}\, +\, [GC_{8}\, +\, BN\, +\, ReLU]_{L5}\, + \,[GC_3]_{L5}
\end{dmath}
% number of parameters: 590'976  % more feat is 16 times more

% concat the results of the corresponding encoder layer after the graph convolution. Similar to ``Jiang et al.''
% \subsubsection*{Full dataset - Equiangular architecture}
% Equiangular sampling, with a latitude bandwidth 384 and longitude bandwidth 576.

@@

\clearpage
\newpage

\begin{description}
	\item[git] \url{https://github.com/nperraud/deepsphere_v2_paper}
	\item[overleaf] \url{https://www.overleaf.com/project/5d31d2fbd6859d54cb707212}
\end{description}

story: there are tradeoffs when designing a spherical CNN -> DS strikes a controllable balance between efficiency / cost and exactitude / equivariance
% DSv2 has a clear tradeoff between equivariance and computational cost

* method development: tool that solve a need (balance btw the desiderata: equiv vs cost)\\
* research question: anisotropy (highlight from the start or leave it as a dangling question in the end)\\

contributions:\\
* theory $\Rightarrow$ convergence (better graph)\\
* experiments on relevant problems (not spherical MNIST) $\Rightarrow$ check the desiderata\\
  * surprising: anisotropy doesn't seem useful\\

\todo{
\begin{itemize}
% 	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN. [Yes]
% 	\item adjacency matrix: $A$ instead of $W$? nath: I am for W
% 	\item anonymize DeepSphere citation - Not doing that
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item citet for inline, citep for parenthesis
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot
	\item result tables: only what supports the main message in the text, everything we have in the appendix
	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, functions. Irregular, uneven. Coarsening, down-sampling.
	\item check English vs American English
	\item unify enumerations. I like (i), (ii), etc.
	\item notation: should we use indices from 1 to n instead of 0 to n-1 for compactness?
	\item add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere.
	\item figure placements
	\item do the experiments on the icosahedral grid with the pooling as defined in the paper
	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
	\item update pygsp and merge sphere branch
	\item check that NN architectures and hyper-params are correctly reported
	\item keep NN architecture and hyper-params in the main text? Move the paragraph to appendix if not enough space.
	%\item shall we still site adam and batch norm? no, mainstream enough
	\item check results for ModelNet40 and integrate in text
	\item Check that HEALPix is introduced in the introduction
	\item heuristic used to set the kernel width in v1. Add values to \figref{kernel_widths} for comparison.
% 	\item check experiments section [Frédérick]: all the results numbers and NN + training details
	\item compute the invariance error of the whole NNs (before and after training, invariance can be learned)
	\item add it with equivariance error of the convolution to \tabref{cosmo}
	\item insert \citet{jiang2019sphericalcnn} in intro: Need a global coordinate system (ok for planets, not projections like cosmo)
\end{itemize}
}

\end{document}

>>>>

\begin{abstract}
% We have to submit soon... The last sentence is a bit disconnected. Yep
Designing a convolution for a spherical neural network requires a delicate tradeoff between efficiency and rotation equivariance. % exactitude with respect to equivariance.
DeepSphere, a method based on a graph representation of the discretized sphere, strikes a controllable balance between these two desiderata.
% Say something here about isotropy?
This contribution is twofold.
First, we study both theoretically and empirically how equivariance is affected by the underlying graph with respect to the number of pixels and neighbors.
% 
Second, we evaluate DeepSphere on relevant problems.
Experiments show state-of-the-art performance and demonstrates the efficiency and flexibility of this formulation.
In particular, comparison with previous work on shape classification and climate event detection suggests that anisotropic filters are an unnecessary price to pay. 
%Say something about speed here... Not sure what?
% memory, computations, implementation complexity
\end{abstract}

@@

\Tabref{shrec17_retrieval} shows the results obtained from the competition's official evaluation script.
% \Tabref{modelnet40} shows results on ModelNet40.

@@

% \begin{table}
%     \centering
%     \begin{tabular}{l cccc}
% 	\toprule
%     & no/no & z/z & SO3/SO3 & z/SO3 \\
% 	\midrule
% 	\citet{cohen2018sphericalcnn} & 85.0 & - & - & - \\
% 	\citet{jiang2019sphericalcnn} & 90.5 & - & - & - \\
% 	\citet{esteves2018sphericalcnn} & - & 88.9 & 86.9 & 76.7 \\
% 	%Esteves \emph{MVCNN} & 94.69 & - & - & - \\
% 	DeepSphere & 87.8 & 86.8 & 86.7 & 76.9 \\
% 	\bottomrule
%     \end{tabular}
%     \caption{
% 		Accuracy results on ModelNet40.
% 		``no'' denotes no augmentation, ``z'' is augmented with rotations around the Z-axis, and ``SO3'' with ZYZ rotations.
% 		\todo{The first is training time augmentation, the second test time augmentation.}
% 	}
%     \label{tab:modelnet40}
% \end{table}

@@

\begin{table}
    \centering
	\begin{tabular}{l l l l l}
		\toprule
        & TC & AR & BG & mean \\
		\midrule
		\cite{mudigonda2017climateevents} & 74 & 65 & 97 & 78.67 \\
		\cite{jiang2019sphericalcnn} (paper) & 94 & 93 & 97 & 94.67 \\
		\cite{jiang2019sphericalcnn} (rerun) & 93.9 & 95.7 & 95.2 & 94.95 \\
        \cite{cohen2019gauge} (S2R) & 97.8 & 97.3 & 97.3 & 97.5 \\
        \cite{cohen2019gauge} (R2R) & 97.9 & 97.8 & 97.4 & 97.7 \\
		\midrule
		DS (Jiang architecture, weighted loss) & 97.1 & 97.6 & 96.5 & 97.1 \\
        DS (weighted loss) & $97.4\pm 1.1$ & $97.7\pm 0.7$ & $98.2\pm 0.5$ & $97.8\pm 0.3$ \\
		DS (wider architecture, weighted loss) & 91.5 & 93.4 & 99.0 & 94.6 \\
		\midrule
        DS (Jiang architecture, non-weighted loss) & 33.6 & 93.6 & 99.3 & 75.5 \\
        DS (non-weighted loss) & $69.2\pm 3.7$ & $94.5\pm 2.9$ & $99.7\pm 0.1$ & $87.8\pm 0.5$ \\
        DS (wider architecture, non-weighted loss) & 73.4 & 92.7 & 99.8 & 88.7 \\
% 		\midrule
% 		\todo{DS (Cohen architecture, weighted loss)} & 95.5 & 96.9 & 93.9 & 95.4 \\
% 		\todo{DS (Cohen architecture, non-weighted loss)} & 56.4 & 90.0 & 98.9 & 81.8 \\
% 		\todo{DS (full dataset, non-weighted loss)} & 31.3 & 75.2 & 99.9 & 68.80 \\
		\bottomrule
    \end{tabular}
    \caption{
		Results on climate event segmentation: accuracy.
		Tropical cyclones (TC) and atmospheric rivers (AR) are the two positive classes, against the background (BG).
		\citet{mudigonda2017climateevents} is not directly comparable as they don't use the same input feature maps.
		Note that a non-weighted cross-entropy loss is not optimal for the accuracy metric.
		\label{tab:climate:accuracy}
	}
\end{table}

@@

\begin{table}
	\centering
	\begin{tabular}{l r r r}
		\toprule
		& \multicolumn{1}{c}{size} & \multicolumn{2}{c}{speed} \\
        \cmidrule(lr){2-2} \cmidrule(lr){3-4}
		& params & inference & training \\
		\midrule
        %Mudigonda et al. & 97 & 74 & 65 & 78.67 & - & - & - \\
		\cite{jiang2019sphericalcnn} & 330\,k & 10\,ms & 10\,h \\ % 328'339
		% \cite{cohen2019gauge} (S2R) & - & - & - \\
        % \cite{cohen2019gauge} (R2R) & - & - & - \\
		DeepSphere (Jiang architecture) & 590\,k & 5\,ms & 3\,h \\ % 590k
        DeepSphere & 13\,M & 33\,ms & 13\,h \\ % 12'926'432
		DeepSphere (wider architecture) & 52\,M & 50\,ms & 20\,h \\
% 		\todo{DS (Cohen architecture)} & 2\,M & 11\,ms & 4\,h \\
% 		\todo{DS (full dataset)} & - & 567\,ms & 48\,h \\
		\bottomrule
    \end{tabular}
    \caption{
		Results on climate event segmentation: size and speed.
	}
		\label{tab:climate:speed}
\end{table}
\pagebreak
% \subsubsection*{DS (Jiang architecture)}

DeepSphere with Jiang architecture\\
encoder:\\
\begin{dmath}
    [GC_{8}\, +\, BN\, +\, ReLU]_{L5}\,+\,\textrm{Pool}\,+\, [GC_{16}\, +\, BN\, +\, ReLU]_{L4}\, +\, \textrm{Pool}\, +\, [GC_{32}\, +\, BN\, +\, ReLU]_{L3}\, +\, \textrm{Pool}\, +\, [GC_{64}\, +\, BN\, +\, ReLU]_{L2}\, +\,\textrm{Pool}\, +\, [GC_{128}\, +\, BN\, +\, ReLU]_{L1}\, +\, \textrm{Pool}\,  +\, [GC_{128}\, +\, BN\, +\, ReLU]_{L0}
\end{dmath}
decoder:\\
\begin{dmath}
    \textrm{Unpool}\, +\,[GC_{128}\, +\, BN\, +\, ReLU]_{L1}\, +\, \textrm{concat}\, +\, [GC_{128}\, +\, BN\, +\, ReLU]_{L1}\, +\, \textrm{Unpool}\, +\, [GC_{64}\, +\, BN\, +\, ReLU]_{L2}\, +\, \textrm{concat}\, +\, [GC_{64}\, +\, BN\, +\, ReLU]_{L2}\, +\, \textrm{Unpool}\, +\, [GC_{32}\, +\, BN\, +\, ReLU]_{L3}\, +\, \textrm{concat}\, +\, [GC_{32}\, +\, BN\, +\, ReLU]_{L3}\, +\,\textrm{Unpool}\, +\, [GC_{16}\, +\, BN\, +\, ReLU]_{L4}\,+\, \textrm{concat}\, +\, [GC_{16}\, +\, BN\, +\, ReLU]_{L4}\, +\,\textrm{Unpool}\,  +\, [GC_{8}\, +\, BN\, +\, ReLU]_{L5}\,+\,\textrm{concat}\, +\, [GC_{8}\, +\, BN\, +\, ReLU]_{L5}\, + \,[GC_3]_{L5}
\end{dmath}
% number of parameters: 590'976  % more feat is 16 times more
Concat is the operation that concatenate the results of the corresponding encoder layer.
% concat the results of the corresponding encoder layer after the graph convolution. Similar to ``Jiang et al.''

@@

% Not similar to Jiang and 4 times the feature maps.

@@

% DeepSphere with a wider architecture is exactly the same, but with 2 times the number of feature maps.

% \subsubsection*{Full dataset - Equiangular architecture}
% Equiangular sampling, with a latitude bandwidth 384 and longitude bandwidth 576.

@@

%\clearpage
%\newpage
%
%\begin{description}
%	\item[git] \url{https://github.com/nperraud/deepsphere_v2_paper}
%	\item[overleaf] \url{https://www.overleaf.com/project/5d31d2fbd6859d54cb707212}
%\end{description}
%
%story: there are tradeoffs when designing a spherical CNN -> DS strikes a controllable balance between efficiency / cost and exactitude / equivariance
%% DSv2 has a clear tradeoff between equivariance and computational cost
%
%* method development: tool that solve a need (balance btw the desiderata: equiv vs cost)\\
%* research question: anisotropy (highlight from the start or leave it as a dangling question in the end)\\
%
%contributions:\\
%* theory $\Rightarrow$ convergence (better graph)\\
%* experiments on relevant problems (not spherical MNIST) $\Rightarrow$ check the desiderata\\
%  * surprising: anisotropy doesn't seem useful\\
%
%\todo{
%\begin{itemize}
%% 	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN. [Yes]
%% 	\item adjacency matrix: $A$ instead of $W$? nath: I am for W
%% 	\item anonymize DeepSphere citation - Not doing that
%% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
%	\item citet for inline, citep for parenthesis
%	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
%	\item PSD of climate data: how does it compare to cosmo and 3D objects?
%	\item align the spheres in figure 1
%	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
%	\item check footnotes
%	\item climate data on spectrum plot
%	\item result tables: only what supports the main message in the text, everything we have in the appendix
%	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, functions. Irregular, uneven. Coarsening, down-sampling.
%	\item check English vs American English
%	\item unify enumerations. I like (i), (ii), etc.
%	\item notation: should we use indices from 1 to n instead of 0 to n-1 for compactness?
%	\item add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere.
%	\item figure placements
%	\item do the experiments on the icosahedral grid with the pooling as defined in the paper
%	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
%	\item update pygsp and merge sphere branch
%	\item check that NN architectures and hyper-params are correctly reported
%	\item keep NN architecture and hyper-params in the main text? Move the paragraph to appendix if not enough space.
%	%\item shall we still site adam and batch norm? no, mainstream enough
%	\item check results for ModelNet40 and integrate in text
%	\item Check that HEALPix is introduced in the introduction
%	\item heuristic used to set the kernel width in v1. Add values to \figref{kernel_widths} for comparison.
%% 	\item check experiments section [Frédérick]: all the results numbers and NN + training details
%	\item compute the invariance error of the whole NNs (before and after training, invariance can be learned)
%	\item add it with equivariance error of the convolution to \tabref{cosmo}
%	\item insert \citet{jiang2019sphericalcnn} in intro: Need a global coordinate system (ok for planets, not projections like cosmo)
%\end{itemize}
%}
%
\end{document}
